### 2017-07-20 paper/paper.tex
Bring things in

% vim: set spell spelllang=en tw=100 :

\documentclass[twoside,11pt]{article}
\usepackage{jair}
\usepackage{theapa}
\usepackage{rawfonts}
\usepackage{complexity}
\usepackage{cleveref}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{nicefrac}
\usepackage{csquotes}

\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}

\jairheading{1}{2016}{0-0}{0/0}{0/0}
\ShortHeadings{Why Real-World Subgraph Matching is Easy}
{McCreesh, Prosser \& Trimble}
\firstpageno{1}

\definecolor{chromaygb1}{rgb}{1.0, 1.0, 0.878431}
\definecolor{chromaygb1b}{rgb}{0.8, 0.972549, 0.72549}
\definecolor{chromaygb2}{rgb}{0.560784,0.929412,0.564706}
\definecolor{chromaygb3}{rgb}{0.407843,0.776471,0.560784}
\definecolor{chromaygb4}{rgb}{0.227451,0.631373,0.552941}
\definecolor{chromaygb5}{rgb}{0.301961,0.431373,0.670588}
\definecolor{chromaygb6}{rgb}{0.317647,0.254902,0.588235}
\definecolor{chromaygb7}{rgb}{0.294118,0.0,0.509804}

\crefname{algorithm}{Algorithm}{Algorithms}
\Crefname{algorithm}{Algorithm}{Algorithms}
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\Crefname{table}{Table}{Tables}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}

\newcommand{\citet}[1]{\citeA{#1}}
\newcommand{\citep}[1]{\cite{#1}}

% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}

\begin{document}

\title{Why Real-World Subgraph Matching is Easy, and \\ What This Means for Filter / Verify in
Graph Databases \thanks{Part of this paper appeared at IJCAI
\citep{DBLP:conf/ijcai/McCreeshPT16}; \cref{section:introduction} has been expanded and
\cref{section:models,section:labelled,section:filterverify} are new material.}}

\author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
           Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \email
           ciaran.mccreesh@glasgow.ac.uk \\
       \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
       \name James Trimble \samethanks \email j.trimble.1@research.gla.ac.uk \\
   \addr University of Glasgow, Glasgow, Scotland}
\maketitle

\begin{abstract}
    Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
    most real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
    this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
    problems, which are computationally challenging with only a couple of hundred vertices in the
    target, and only twenty to thirty pattern vertices. For the non-induced version of the problem,
    these instances lie on a satisfiable / unsatisfiable phase transition, whose location we can
    predict; for the induced variant, much richer behaviour is observed, and constrainedness gives a
    better measure of difficulty than does proximity to a phase transition. In both cases, we can
    use this to understand and improve variable and value ordering heuristics.

    When labels are introduced, as commonly seen in real-world applications and in graph databases,
    additional rich behaviour arises.  We show that when even the simplest constraint programming
    algorithms are used, all commonly used benchmark suites for evaluating graph database systems
    become trivial, and that existing notions of difficulty are based almost entirely upon
    unnecessarily poor performance from overly simplistic subgraph isomorphism algorithms. With this
    in mind, we demonstrate that the widely researched ``filter~/ verify'' indexing technique used in
    graph databases is a flawed design based upon a repeated misunderstanding of what makes subgraph
    isomorphism hard. We explain why filter~/ verify cannot be beneficial, even with harder
    instances than those currently used, when paired with any reasonable modern subgraph isomorphism
    algorithm.
\end{abstract}

\section{Introduction}\label{section:introduction}

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
vertices of a given pattern graph to the vertices of a given target graph which preserves
adjacency---in essence, we are ``finding a copy of'' the pattern inside the target. The
\emph{induced} variant of the problem additionally requires that the mapping preserve non-adjacency,
so there are no ``extra edges'' in the copy of the pattern that we find. We illustrate both variants
in \cref{figure:sip}.  Although these problems are \NP-complete, modern practical subgraph
isomorphism algorithms can handle problem instances with many hundreds of vertices in the pattern
graph, and up to ten thousand vertices in the target graph
\citep{DBLP:journals/pami/CordellaFSV04,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
leading to successful applications in areas such as computer vision
\citep{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, biochemistry
\citep{o:10.1371/journal.pone.0076911,DBLP:conf/gbrpr/CarlettiFV15}, and pattern recognition
\citep{DBLP:journals/ijprai/ConteFSV04}.  In particular, a labelled version of subgraph isomorphism
is one of the key components in supporting complex queries in graph databases---typically, these
systems store a fixed set of target graphs, and for a sequence of pattern queries, they must return
every target graph which contains that pattern.  One common approach to this problem is to combine a
subgraph isomorphism algorithm with an indexing system in a so-called ``filter~/ verify'' model,
where invariants are used to attempt to pre-exclude unsatisfiable instances to avoid the cost of a
subgraph isomorphism call. In this context, the terms \emph{matching} and \emph{verification} are
often used for the subgraph isomorphism step (or a slightly broader problem, for example permitting
wildcards).

However, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances involving hundreds
or thousands of vertices. Experimental evaluations of subgraph isomorphism algorithms are usually
performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
and graphs that encode biochemistry and computer vision problems. Using random instances to evaluate
algorithm behaviour can be beneficial, because it provides a way of generating many instances
cheaply, and reduces the risk of over-fitting when tuning design parameters. The random instances
used typically come from common datasets
\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh)
graph and permuting the vertices. Such instances are guaranteed to be
satisfiable---\citet{o:AntonO09} exploited this property to create large sets of random satisfiable
boolean satisfiability instances.  This is the most common approach to generating random subgraph
isomorphism instances, meaning existing benchmark suites contain relatively few non-trivial
unsatisfiable instances (although a few of the patterns in the instances by
\citeauthor{DBLP:journals/constraints/ZampelliDS10} have had extra edges added, making them
unsatisfiable). Also, the satisfiable instances tend to be computationally fairly easy, with most of
the difficulty being in dealing with the size of the model. This has lead to bias in algorithm
design, to the extent that some proposed techniques, such as those of
\citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.

The first contribution of this paper is to present and evaluate new methods for creating random
pattern~/ target pairs.  The method we introduce generates both satisfiable and unsatisfiable
instances, and can produce computationally challenging instances with only a few tens of vertices in
the pattern, and 150 vertices in the target. This is not entirely straightforward---the lack of
unsatisfiable instances for testing purposes cannot be addressed simply by taking a pattern graph
from one of the existing random suites with the ``wrong'' target graph, as this tends to give either
a trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
case that a relatively small random graph is unlikely to appear in a larger random graph.)

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.3,every node/.style={font=\footnotesize}]
        \begin{scope}
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};

            \draw [ultra thick] (Na) -- (Nb);
            \draw [ultra thick] (Nb) -- (Nc);
            \draw [ultra thick] (Nc) -- (Nd);
            \draw [ultra thick] (Nb) -- (Nd);

            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};

            \draw [] (N1) -- (N2);
            \draw [ultra thick] (N1) -- (N3);
            \draw [] (N1) -- (N4);
            \draw [] (N2) -- (N4);
            \draw [ultra thick] (N3) -- (N5);
            \draw [ultra thick] (N3) -- (N6);
            \draw [] (N4) -- (N6);
            \draw [ultra thick] (N5) -- (N6);
            \draw [] (N2) to [in=45, out=315] (N6);
            \draw [] (N1) -- (N7);
            \draw [] (N5) -- (N9);
            \draw [] (N7) -- (N8);
            \draw [] (N8) -- (N9);
        \end{scope}
        \begin{scope}[xshift=20cm]
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};

            \draw [ultra thick] (Na) -- (Nb);
            \draw [ultra thick] (Nb) -- (Nc);
            \draw [ultra thick] (Nc) -- (Nd);
            \draw [ultra thick] (Nb) -- (Nd);

            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};

            \draw [ultra thick] (N1) -- (N2);
            \draw [] (N1) -- (N3);
            \draw [ultra thick] (N1) -- (N4);
            \draw [ultra thick] (N2) -- (N4);
            \draw [] (N3) -- (N5);
            \draw [] (N3) -- (N6);
            \draw [ultra thick] (N4) -- (N6);
            \draw [] (N5) -- (N6);
            \draw [ultra thick, dashed] (N2) to [in=45, out=315] (N6);
            \draw [] (N1) -- (N7);
            \draw [] (N5) -- (N9);
            \draw [] (N7) -- (N8);
            \draw [] (N8) -- (N9);
        \end{scope}
    \end{tikzpicture}

    \caption{On the left, an induced subgraph isomorphism. On the right, a non-induced subgraph
    isomorphism: the extra dashed edge is not present in the pattern graph.}
    \label{figure:sip}
\end{figure}

This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism we identify three relevant control
parameters: we can independently alter the edge probability of the pattern graph, the edge
probability of the target graph, and the relative orders (number of vertices) of the pattern and
target graphs.  For non-induced isomorphisms, with the correct choice of parameters we see results
very similar to those observed with boolean satisfiability problems: there is a phase transition
(whose location we can predict) from satisfiable to unsatisfiable, and we see a solver-independent
complexity peak near this phase transition. Additionally, understanding this behaviour
helps us to select better variable- and value-ordering heuristics---this is the second contribution
of this paper.

For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be hard, despite it not being near a phase transition. To show that this is
not a simple weakness of current subgraph isomorphism algorithms, we verify that this region is also
hard when using a pseudo-boolean encoding, and under reduction to the clique problem. Interestingly,
the constrainedness measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult
region---the third contribution of this paper is to use these instances to provide evidence in
favour of constrainedness, rather than proximity to a phase transition, being an accurate predictor
of difficulty, and to show that constrainedness is not simply a refinement of a phase transition
prediction.

% What about other random graph models? Graphs where every vertex has the same degree, which are known
% as $k$-regular, exhibit very similar characteristics.

When labels on vertices are introduced, as is commonly seen in real-world applications and in graph
database systems, richer behaviour emerges, particularly when moving away from a uniform labelling
scheme, and we see that VF2 behaves much worse than the Glasgow and LAD algorithms in certain cases.
A close look shows that these cases \emph{should} be easy to solve.

This is not simply a theoretical curiosity. A labelled version of subgraph isomorphism is one of
the key components in supporting complex queries in graph databases---typically, these systems store
a fixed set of target graphs, and for a sequence of pattern queries, they must return every target
graph which contains that pattern.  One common approach to this problem is to combine a subgraph
isomorphism algorithm with an indexing system in a so-called ``filter~/ verify'' model, where
invariants are used to attempt to pre-exclude unsatisfiable instances to avoid the cost of a
subgraph isomorphism call. In this context, the terms \emph{matching} and \emph{verification} are
often used for the subgraph isomorphism step (or a slightly broader problem, for example permitting
wildcards).

We look at some of the datasets commonly used to test graph database systems, which the literature
suggests are extremely hard to solve. Simple experiments show that the entire perceived difficulty
of each of these datasets comes down to the widespread use of especially poor subgraph isomorphism
algorithms. Our fourth contribution is to show that the use of forward-checking and the right
variable ordering heuristic, even without the more sophisticated recent advances in subgraph
algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally, we
explain why filter~/ verify (and other recently proposed techniques) cannot be beneficial even on
more challenging instances except if an extremely poor choice of subgraph isomorphism algorithm is
made.

\section{Experimental Setup}

The experiments in this paper are performed on systems with Intel Xeon E5-4650 v2 CPUs, running
Scientific Linux release 6.7. We will be working with three subgraph isomorphism solvers: the
Glasgow solver \citet{DBLP:conf/cp/McCreeshP15}; LAD \citep{DBLP:journals/ai/Solnon10}; and VF2
\citep{DBLP:journals/pami/CordellaFSV04}. Each was compiled using GCC 4.9.

Recall that the Glasgow and LAD solvers use backtracking search to build up an assignment of target
vertices (values) to pattern vertices (variables), but differ in terms of inference and ordering
heuristics.  The approach used by VF2 is similar, although the domains of variables are not stored
(in the style of conventional backtracking, rather than forward-checking), and so domain wipeouts
are not detected until an assignment is made. Both the Glasgow and LAD solvers use domain sizes to
guide search, whilst VF2 uses adjacency to previously-assigned vertices.

We measure only the number of recursive calls (guessed assignments) made, not
runtimes. We are not aiming to compare absolute performance between solvers; rather, we are looking
for solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which
is enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
for extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
with this timeout, so our picture of just how hard the hardest instances are with these solvers is
less detailed.

\section{Non-Induced Subgraph Isomorphisms}

\begin{figure}[t]
    \centering
    \includegraphics*{plots/phase-transition.pdf}

    \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
        probability of 0.40, and varying pattern edge probability, we observe a phase transition and
        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
        one instance. The lines show mean search effort and mean proportion satisfiable.}
    \label{figure:phase-transition}
\end{figure}

Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, a fixed
target edge probability of 0.40, and no vertex labels. As we vary the pattern edge probability from
0 to 1, we would expect to see a shift from entirely satisfiable instances (with no edges in the
pattern, we can always find a match) to entirely unsatisfiable instances (a maximum clique in this
order and edge probability of target graph will usually have between 9 and 12 vertices). The move
from green circles (satisfiable) to blue crosses (unsatisfiable) in \cref{figure:phase-transition}
shows that this is the case. For densities of 0.67 or greater, no instance is satisfiable; with
densities of 0.44 or less, every instance is satisfiable; and with a density of 0.55, roughly half
the instances are satisfiable.

The line plots mean search effort using the Glasgow solver: for sparse patterns, the problem is
trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
complexity peak around the point where half the instances are satisfiable.  We also plot the search
cost of individual instances, as points. The behaviour we observe looks remarkably similar to the
clique decision problem and to random 3SAT problems---compare, for example,
Figure~1 of
\citet{DBLP:journals/cacm/Leyton-BrownHHX14}. In particular, satisfiable instances tend to be easier, but show greater
variation than unsatisfiable instances, and there are exceptionally hard satisfiable instances
\citep{DBLP:conf/cp/SmithG97}. (The Glasgow solver supports
parallel search with a work-stealing strategy explicitly designed to reduce or eliminate these. We
are not using parallel search to avoid dealing with the complexity of search tree
measurements under parallelism.)

\begin{figure}[tb]
    \centering
    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
        \matrix {
            \node [anchor=center] {}; &
            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
            \\
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-30-150.pdf}}; \\
        };
    \end{tikzpicture}

    \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
        pattern edge probability and the y-axis is the target edge probability, both from 0 to 1.
        Along the top row, we show the proportion of instances which are satisfiable; the white
        bands show the phase transitions, and the black lines are the predictions using
        equation~\eqref{equation:non-induced-prediction} of where the
        phase transition will occur. On the subsequent three rows, we show the number of search nodes used by the
        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances.}
    \label{figure:non-induced}
\end{figure}

What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis) and
target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
of 0.01. For each of these points, we generate ten random instances. The colour denotes the
proportion of these instances which were found to be satisfiable.  Inside the red region, at the
bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
pattern in a sparse target. In the green region, at the top left, every instance is
satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
have to preserve adjacency, not non-adjacency). The white band between the regions shows the
location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
black line below.)

On subsequent rows, we show the average number of search nodes used by the different algorithms. In
general, satisfiable instances are easy, until very close to the phase transition. As we hit the
phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as we
pass through the phase transition and move deeper into the unsatisfiable region, instances become
easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
than Glasgow or LAD. VF2 also finds some instances extremely hard when the target is empty but the
pattern is not---this turns out to be extremely important, and we return to it in
\cref{section:labelled,section:filterverify}. Thus, although we have moved away from a single
control parameter, we still observe the easy-hard-easy pattern seen in many \NP-complete
problems.

\subsection{Locating the Phase Transition}

We can approximately predict the location of the phase transition by calculating (with
simplifications regarding rounding and independence) the expected number of solutions for given
parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
p + 1) \] possible assignments of target vertices to pattern vertices.  If we assume the pattern
has exactly $d_p \cdot \binom{p}{2}$ edges, we obtain the probability of all of these edges being mapped to
edges in the target by raising $d_t$ to this power, giving an expected number of solutions of
\begin{equation}\label{equation:non-induced-prediction}
    \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.}
\end{equation} This
formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
in the first row of \cref{figure:non-induced}.

This prediction is generally reasonably accurate, except that for very low and very high pattern
densities, we overestimate the satisfiable region. This is due to variance: although an expected
number of solutions much below one implies a high likelihood of unsatisfiability, it is not true
that a high expected number of solutions implies that any particular instance is likely to be
satisfiable. (Consider, for example, a sparse graph which has several isolated vertices. If one
solution exists, other symmetric solutions can be obtained by permuting the isolated vertices.
Thus although the expected number of solutions may be one, there cannot be exactly one solution.) A
similar behaviour is seen with random constraint satisfaction problems
\citep{DBLP:journals/ai/SmithD96}.

\subsection{Variable and Value Ordering Heuristics}

Various general principles have been considered when designing variable and value ordering
heuristics for backtracking search algorithms---one of these is to try to maximise the expected
number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.  This is
usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD and
Glasgow pick a variable with fewest remaining values in its domain: doing this will generally reduce
the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or more
domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick a
variable corresponding to a pattern vertex of highest degree. This strategy was determined
empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
exponent in the second half of the formula, maximising the overall value. LAD does not apply a value
ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.  Again, this was
determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing the
remaining target density. The VF2 heuristics, in contrast, are based around preserving connectivity,
which gives very little discrimination except on the sparsest of inputs.

\section{Induced Subgraph Isomorphisms}

\begin{figure}[p]
    \centering
    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
        \matrix {
            \node [anchor=center] {}; &
            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(15, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-30-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-30-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-30-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-30-150.pdf}}; \\
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Constrainedness}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-30-150.pdf}}; \\
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Complement?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-14-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-15-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-16-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-30-150.pdf}}; \\
        };
    \end{tikzpicture}
    \caption{Behaviour of algorithms on the induced variant, shown in the style
    of \cref{figure:non-induced}. The fifth row plots constrainedness using
    equation~\eqref{equation:constrainedness}: the
    darkest region is where $\kappa = 1$, and the lighter regions show where the problem is either
    over- or under-constrained. The final row shows when the Glasgow algorithm performs better when
    given the complements of the pattern and target graphs as inputs---the solid lines show the
    location of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
\end{figure}

In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
half of the plots resemble the non-induced results, and the top left half is close to a mirror
image. The central satisfiable region, which is away from either phase transition, is
computationally easy, but instances near the phase transition are hard.

For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
not being near either phase transition, instances in the centre remain computationally challenging.
We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.

We might expect these complexity plots to be symmetric along the diagonal, since for the induced
problem, if we replace both inputs with their complements, the solutions remain the same.  For the
Glasgow solver, this is the case. This should be expected, because this complement property is
precisely how the Glasgow solver handles the induced variant (although the heuristics may differ
between the two, which we discuss below). For LAD, some of the very dense patterns are slightly
harder than their diagonal opposites (LAD reasons about degrees, but not about complement-degrees);
it is interesting to note that for larger target graphs, VF2 finds \emph{all} dense pattern graphs
difficult.

\subsection{Predictions and Heuristics}

To predict the location of the induced phase transition, we repeat the argument for locating the
non-induced phase transition and additionally consider non-edges, to get an expected number of
solutions of \begin{equation}\label{equation:induced-prediction} \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \end{equation} We plot this using black lines on
the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
very dense patterns.

We might guess that degree-based heuristics would just not work for the induced problem: for any
claim about the degree, the opposite will hold for the complement constraints. However, empirically,
this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
(The only steps performed by the Glagsow algorithm which differ under taking the complements are
the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
step using degree information, but does not consider the complement degree, and VF2 uses
connectivity in the pattern graph.)

For patterns of order 10, it is always better to try to move towards the satisfiable region: if we
are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
towards the top left), and if we are in the top left we should use the complement instead. This
goes against a suggestion by \citet{DBLP:conf/aaai/Walsh98} that switching heuristics based upon an estimate of
the solubility of the problem may offer good performance.

For larger patterns, more complex behaviour emerges. If we are in the intersection of the bottom half
and the bottom right diagonal of the search space, we should always retain the original heuristic,
and if we are in the intersection of the top half and the top left diagonal, we should always use
the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
use, and using directional derivatives is not enough to resolve the problem. A close observation of
the data suggests that the actual location of the phase transition may be involved (and perhaps
\citeauthor{DBLP:conf/aaai/Walsh98}'s suggestion applies only in these conditions). In any case, $\langle Sol
\rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.

In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
are usually easy. In this situation, these experiments justify reusing the
non-induced heuristics on induced problems.

\subsection{Is the Central Region Genuinely Hard?}

\begin{figure}[p]
    \centering
    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
        \matrix {
            \node [anchor=center] {}; &
            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(12, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(18, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $G(25, x)$\hspace*{3em}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-25-75.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-25-75.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Clasp (PB)}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-25-75.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glucose (SAT)}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-25-75.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-25-75.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize BBMC (Clique)}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-10-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-12-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-14-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-16-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-18-75.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-25-75.pdf}}; &
            \\
        };
    \end{tikzpicture}
    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
        \cref{figure:non-induced}. The second row shows the number of search nodes used by the
    Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
and SAT solvers, the fifth shows the number of search nodes used on the MIP encoding, and the
final row the clique encoding.}\label{figure:alt}
\end{figure}

The region in the parameter space where both pattern and target have medium density is far from a
phase transition, but nevertheless contains instances that are hard for all three solvers. We would
like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot reason
about adjacency and non-adjacency simultaneously?), or whether instances in this region are
inherently difficult to solve.  Thus we repeat the induced experiments on smaller pattern and target
graphs, using different solving techniques.  Although these techniques are not competitive in
absolute terms, we wish to see if the same pattern of behaviour occurs. The results are plotted in
\cref{figure:alt}.

Our pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
$w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
Constraints are added to ensure that each pattern vertex maps to exactly one target vertex, that
each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are mapped to
adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We use the
Clasp solver \citep{DBLP:journals/aicom/GebserKKOSS11} version 3.1.3 to solve the pseudo-boolean instances.  The
instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
inside the central region, and the easy satisfiable instances remain easy. Similar results are seen
with the Glucose SAT solver \citep{o:glucose} using a direct encoding of the cardinality constraints.
We also show an integer program encoding: the Gurobi solver is only able to solve some of the
trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
limit.

The \emph{association graph encoding} of a subgraph isomorphism problem is a reduction to the clique
decision problem. \citet{DBLP:conf/cp/McCreeshNPS16} describe and study this approach in more
detail. Briefly, the association graph is constructed by creating a new graph with a vertex for each
pair $(p, t)$ of vertices from the pattern and target graphs respectively. There is an edge between
vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$ to $t_1$ and $p_2$ to $t_2$
simultaneously is permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and only if $t_1$ is adjacent to
$t_2$. A clique of size equal to the order of the pattern graph exists in the association graph if
and only if the problem is satisfiable \citep{o:Levi73}. We used this encoding with an
implementation of \citet{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm,
modified to solve the decision problem rather than the optimisation problem.  Again, our results
show that the instances in the central region remain hard, and additionally, some of the easy
unsatisfiable instances become hard.

Together, these experiments suggest that the central region may be genuinely hard, despite not being
near a phase transition. The clique results in particular rule out the hypothesis that subgraph
isomorphism solvers only find this region hard due to not reasoning simultaneously about adjacency
and non-adjacency, since the constraints in the association graph encoding consider compatibility
rather than adjacency and non-adjacency.

\subsection{Constrainedness}

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be underconstrained, and is
likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
is very small or very large are usually easy. By handling injectivity as a restriction on the size
of the state space rather than as a constraint, we derive
\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left(
t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
\binom{p}{2}} \right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the
fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition---our method generates instances where constrainedness and ``close to a phase
transition'' give very different predictions, and constrainedness gives the better prediction.

Unfortunately, constrainedness does not help us with heuristics: minimising constrainedness gives
the same predictions as maximising the expected number of solutions.

\section{$k$-Regular Graphs}\label{section:models}

\begin{figure}[tb]
    \centering
    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
        \matrix {
            \node [anchor=center] {}; &
            \node [anchor=center] {\scriptsize $R(10, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $R(20, x)$\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize $R(30, x)$\hspace*{3em}}; &
            \\
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-30-150.pdf}}; &
            \\[0.3cm]
            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-30-150.pdf}}; &
            \\
        };
    \end{tikzpicture}

    \caption{Behaviour of algorithms on random $k$-regular graphs, as the pattern degree (x-axis)
    and target degree (y-axis) are varied from 0 to $\left|\operatorname{V}\right| - 1$.}
    \label{figure:kr}
\end{figure}

What about richer graph structures? We now look at what happens with the non-induced problem if we
use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then see the effects of
introducing vertex labels. Both models still exhibit phase transition behaviour, although with
sufficiently many labels the ``satisfiable'' region now contains a mix of satisfiable and
unsatisfiable instances.

?? We then move on to more structured instances.

By $R(n, k)$ we mean a random graph with $k$ vertices, each of which has degree $k$. In
\cref{figure:kr} we recreate \cref{figure:non-induced}, using the NetworkX implementation
\citep{o:NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
The $x$-axes range from degree 0 to degree $n - 1$, and the y-axes range from degree 0 to degree
149.  The satisfiable / unsatisfiable plots show very similar results to the non-induced problem on
Erd\H{o}s-R\'enyi random graphs, with a similarly sharp phase transition, and
again only instances near the phase transition are difficult.

?? Talk about degree heuristics

\section{Labelled Graphs}\label{section:labelled}

So far, we have looked at unlabelled graphs. What happens when labels on vertices are introduced?
This is common in real-world applications---for example, when working with graphs representing
chemical molecules, mappings are typically expected only to match carbon atoms with carbon atoms,
hydrogen atoms with hydrogen atoms, and so on.  We will look at the non-induced variant, as this
seems to be more common in the literature.
% , and return to the Erd\H{o}s-R\'enyi model.

\subsection{Predictions and Empirical Hardness}

Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
small compared to $p$. Let $\ell(v)$ be the label of vertex $v$. By defining \[
    \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = n \} \] to be the set of
vertices with label $k$, we can partition the pattern vertices by label into disjoint sets \[\{
    \operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of
which is expected to contain $\nicefrac{p}{k}$ vertices. Similarly, we may partition the target
vertices into disjoint sets $\{ \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.

Without labels, there are $t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)$
possible injective assignments of target vertices to pattern vertices.  With labels, observe that
for any label $x$, vertices in $\operatorname{V}(P)|_x$ may only be mapped to vertices in
$\operatorname{V}(T)|_x$.  Thus for each label $x$, we have an expected $\nicefrac{p}{k}$ variables,
each of whose domains contain $\nicefrac{t}{k}$ values. We would like to say that the size of the
state space is now \[ |S| = \left((\nicefrac{t}{k})^{\underline{\nicefrac{p}{k}}}\right)^{k}
\textnormal{,} \] but to do this we must state what $a^{\underline{b}}$ means when $b$ is
fractional. The gamma function $\Gamma(n)$ is equal to $(n - 1)!$ for integers $n \ge 1$, but is also
defined for positive real numbers, obeying the identify $\Gamma(x + 1) = x\Gamma(x)$. By noting that
$t^{\underline{p}} = \nicefrac{t!}{(t - p)!}$, we may obtain a reasonable
continuous extension by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k} +
1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,,} \]

As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, and so if we simplify by
assuming the pattern will have \emph{exactly} this many edges, we obtain the
probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
power, giving an estimate of \begin{equation}\label{equation:non-induced-labelled-prediction} \langle Sol \rangle = \left(
    \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
1\right)} \right)^{k}  \cdot
{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \end{equation}

\begin{figure}[t]
    \centering
    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
        \matrix {
            \node [anchor=center] {}; &
            \node [anchor=center] {\scriptsize No labels\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize 2 labels\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize 3 labels\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize 5 labels\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize 10 labels\hspace*{1em}}; &
            \node [anchor=center] {\scriptsize 20 labels\hspace*{2em}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l2-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l3-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l5-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l20-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l2-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l3-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l5-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l20-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l2-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l3-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l5-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l20-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l2-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l3-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l5-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l20-150.pdf}}; &
            \\
            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2 Failures}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l2-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l3-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l5-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l10-150.pdf}}; &
            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l20-150.pdf}}; &
            \\
        };
    \end{tikzpicture}
    \caption{On the top row, predicted and actual location of the phase transition for labelled
    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 150,
    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On
    subsequent rows, the average number of search nodes needed to solve an instance, for three
    different solvers, and on the final row, the proportion of instances aborted due to a timeout
    with VF2.}\label{figure:labels}
\end{figure}

So how good are these predictions? The black lines on the first row of heatmaps in
\cref{figure:labels} plot where we calculate $\langle Sol \rangle = 1$ will occur. For small numbers
of labels, our predictions are slightly better than in the unlabelled case: there seems to be less
of a variance problem for very dense patterns. Even as the numbers of labels becomes relatively
large, the prediction of the phase transition still occurs in the right place, but with ten labels
we start to see sporadic unsatisfiable instances deep inside the satisfiable region. With twenty
labels, we instead get a kind of phase transition from ``all unsatisfiable'' to ``mixed satisfiable
and unsatisfiable''. We can understand this intuitively: with sufficiently many labels, we might
generate, say, a red vertex adjacent to a blue vertex in the pattern, but not in the target.  With
twenty labels, we even sometimes generate no vertices with a particular label in the target at all.
This in many ways resembles ``flaws'' generated by certain random constraint satisfaction problem
instance generators
\citep{DBLP:journals/constraints/AchlioptasMKSKK01,DBLP:journals/constraints/GentMPSW01}.

What about empirical hardness? As before, some instances on the phase transition are hard for all
solvers (although as the number of labels increases, the hardest instances become easier).
Instances far from the phase transition are easy, except for VF2, which occasionally finds
some of instances with larger numbers of labels very difficult. This behaviour occurs both on
satisfiable instances, and on the flawed unsatisfiable instances deep inside the ``satisfiable''
region.  It is interesting to observe that all such unsatisfiable instances that VF2 finds difficult
have a very small proof of unsatisfiability, with neither Glasgow nor LAD requiring more than one
hundred recursive calls.

\subsection{Richer Label Models, and VF2's Deficiencies}\label{subsection:richer-labels}

Why does VF2 find some of these instances so hard? VF2 does not track domains, like other
algorithms, and does not detect a domain wipeout until it branches on a particular vertex.  It also
cannot detect small domains, and only uses ``adjacency to an existing assignment'' as a branching
heuristic. This can make it very hard for VF2 to detect that it is in an obviously failed state, or
that an instance or subproblem is trivially unsatisfiable.

\begin{figure}[t]
    \centering
    \includegraphics*{plots/skewed.pdf}

    \caption{Cumulative number of instances solved over time, using the richer model of randomness
    described in \cref{subsection:richer-labels}.}
    \label{figure:skewed-cumulative}
\end{figure}

To illustrate this point further, we will now look at some instances created using a slightly more
structured random model. We create a family of one thousand labelled instance pairs, as follows. To
create a pattern graph, we create ten vertices with label zero, with edges between these vertices
with probability 0.2. We then add another ten vertices, with labels chosen randomly between one and
thirty, and add edges between these vertices and the zero-labelled vertices with probability 0.1.
To generate a target, we follow a similar process: we have fifty vertices with label zero and edge
probability 0.2, fifty vertices with labels randomly between one and thirty, and edges from the
first set of vertices to the second set with probability 0.3.  This model was selected and the
parameters tuned to provide a demonstration of particular behaviours of VF2, not because of any natural
property (although inspiration came from seeking a very crude approximation of chemical graphs,
which often contain a lot of carbon in the centre, and other atoms around the outside). From our set
of one thousand such instances, sixty six are satisfiable.

We plot the cumulative number of instances solved over time for these instances in
\cref{figure:skewed-cumulative}. Both Glasgow and LAD find all of these instances trivial, but VF2
finds many of them extremely difficult.  The ``VF2${\leftrightarrow}$'' line shows what happens with
VF2 if the graphs are permuted, so that the non-zero labelled instances are given lower vertex
numbers rather than higher vertex numbers. This
makes VF2 behave better, but still extremely poorly compared to the other two solvers; we return to
this in reference to work by \citet{DBLP:conf/edbt/KatsarouNT17} in the following section.
It is also important to note for later that both VF2 variations find some satisfiable and some
unsatisfiable instances extremely hard.

\begin{figure}[tb]
    \centering
    \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
        \begin{scope}
            \newcount \myc
            \newcount \myd
            \newcount \mye
            \foreach \n in {1, ..., 9}{
                \myc=\n \advance\myc by -1 \multiply\myc by -360 \divide\myc by 10 \advance\myc by 18.0
                \node[draw, circle, fill=chromaygb1, inner sep=0.5pt] (N\n) at (\the\myc:2.4) {\phantom{0}};
            }
            \node[draw, circle, fill=chromaygb3, inner sep=0.5pt] (N10) at (54.0:2.4) {\phantom{0}};

            \foreach \n in {1, ..., 8}{
                \myd=\n
                \myc=\n \advance\myc by 1
                \foreach \m in {\the\myc, ..., 9}{
                    \mye=\m
                    \draw (N\the\myd) -- (N\the\mye);
                }
            }
        \end{scope}
        \begin{scope}[xshift=9cm]
            \newcount \myc
            \newcount \myd
            \newcount \mye
            \foreach \n in {1, ..., 9}{
                \myc=\n \advance\myc by -1 \multiply\myc by -360 \divide\myc by 10 \advance\myc by 18.0
                \node[draw, circle, fill=chromaygb1, inner sep=0.5pt] (N\n) at (\the\myc:2.4) {\phantom{0}};
            }
            \node[draw, circle, fill=chromaygb7, inner sep=0.5pt] (N10) at (54.0:2.4) {\phantom{0}};

            \foreach \n in {1, ..., 8}{
                \myd=\n
                \myc=\n \advance\myc by 1
                \foreach \m in {\the\myc, ..., 9}{
                    \mye=\m
                    \draw (N\the\myd) -- (N\the\mye);
                }
            }
        \end{scope}
    \end{tikzpicture}
    \caption{A trivially unsatisfiable subgraph isomorphism instance which VF2 finds exponentially
    difficult. The pattern (left) and target (right) both consist of a clique, plus one extra vertex
    which has a different label in each graph.}\label{figure:vf2-stupidity}
\end{figure}

An even more extreme example of VF2's misbehaviour can be seen in \cref{figure:vf2-stupidity}. Here
we have a pattern graph and a target graph, both of which are cliques plus one isolated vertex, and
the isolated vertices have different unique labels. This is trivially unsatisfiable (and Glasgow and
LAD detect this without search), but unless the labelled vertex is given the lowest vertex number
(so it is branched on first), VF2 takes nearly a million recursive calls to detect this: VF2 will
try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
This example can also be extended slightly to fool any simple static heuristic, or any simple label
counting mechanism.

These experiments further highlight that VF2 occasionally finds some instances which should be easy hard.
But does this cause problems in practice? The literature suggests that it does. For example,
\citet{o:Gromping14} uses VF2 in a package for the R statistics language, and states
\begin{displayquote}``There are (not so many) instances for which creation of a clear design is
    prohibitively slow in the current implementation that evaluates subgraph isomorphism with the
    VF2 algorithm \ldots Recent experiences with a few of these showed that the LAD algorithm was
    very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
Similarly, \citet{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
(and prohibitively high) compile times in some cases due to the expense of subgraph isomorphism
calls---given the heavily labelled nature of these graphs, we conjecture that any domain-based
algorithm would eliminate this cost.

However, by far the biggest problem is in graph databases. The following section shows how
widespread use of VF2 has not just lead to overly pessimistic conclusions regarding performance, but
has misdirected the design of larger systems.

\section{Querying Graph Databases}\label{section:filterverify}

A particularly common use of subgraph isomorphism algorithms is inside graph databases.  The
general problem these systems solve is, for a set of target graphs, to process a pattern query and
return every target graph which is subgraph-isomorphic to that pattern. The set of target graphs is
usually seen as fixed, or at least rarely-changing, whilst the patterns arrive dynamically. This has
lead to the development of systems which perform extensive computations on the target graphs
beforehand, in the hopes of reducing the response times for individual pattern queries. The most
popular of these strategies is a form of indexing which is often named \emph{filter~/ verify}.

\subsection{The Filter / Verify Paradigm}

The filter~/ verify approach has an interesting history, of which we now give a very selective and
incomplete overview. Our description is biased by a general modern understanding of the empirical
hardness of \NP-complete problems, which was not widely known at least at the time of the earlier
papers we discuss. The common theme of all of the following papers is that pre-computed information
is used to eliminate certain unsatisfiable instances from consideration, without performing a
subgraph isomorphism test. For example, an index might contain a bit of information expressing
whether a target graph contains two red vertices. When a pattern graph with two red vertices is used
as a query, any target whose feature set does not have this bit set would not be considered, and so
a subgraph isomorphism call would not be made for that pattern / target pair. (In practice, feature
hashing is often used, which can lead to false positives, but this is not relevant to our
discussion.)

An early graph database system by \citet{DBLP:conf/pods/ShashaWG02} uses a filtering heuristic to
eliminate unsatisfiable instances based upon simple structural elements. It is not clear whether the
aim is to reduce I/O costs, or to reduce the number of queries which must be tested, and the
experiments do not answer whether the indexing is effective. However, the work was influenced by a commercial
graph database system, whose documentation \citep{o:Daylight} states that indexing is
used to minimise disk accesses.

Subsequently, in a widely cited paper, \citet{DBLP:conf/sigmod/YanYH04} introduce an indexing system
called gIndex. Again, this system handles queries by first producing a set of candidates by
eliminating certain unsatisfiable instances, this time by using substructures. They argue that the
query response time, which is to be minimised, is governed by the equation \[ T =
T_{\mathit{search}} + \left|C_q\right| \cdot T_{\mathit{iso\_test}} \textnormal{,}\] where
$T_{\mathit{search}}$ is the time taken to search for a candidate set of potential solutions of size
$\left|C_q\right|$, and $T_{\mathit{iso\_test}}$ is the cost of a subgraph isomorphism test. They
reason that since isomorphism testing is \NP-complete, by making $\left|C_q\right|$ as small as
possible, the query response time will be reduced.  (Importantly, it is \emph{not} time taken to
load graphs from disk which contributes to the per-candidate cost, but rather the time to perform a
subgraph isomorphism call.) They conclude that ``graph indexing plays a critical role at efficient
query processing in graph databases''.

This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
separate I/O and isomorphism testing costs, obtaining a query response time of \[
    T_{\mathit{search}} + \left|C_q\right| \cdot (T_{\mathit{io}} + T_{\mathit{iso\_test}})
    \textnormal{.} \] The authors explicitly state that ``the value of $T_{\mathit{iso\_test}}$ does
not change much for a given query''. The argument presented is that
\begin{displayquote}``Sequential scan is very costly
because one has to not only access the whole graph database but also check subgraph isomorphism. It
is known that subgraph isomorphism is an \NP-complete problem. Clearly, it is necessary to build
graph indices in order to help processing graph queries.''\end{displayquote}
With what we now know about the behaviour of modern subgraph isomorphism algorithms, and the nature
of solving \NP-complete problems in general, we should immediately be suspicious of these claims. We
do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
input graphs are similar. In particular, any instance which can be excluded based upon filtering
must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
decent matching algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
should make us wonder whether the subgraph isomorphism algorithm being used was excessively
primitive. Indeed, the matching algorithm used is described only as ``the simplest approach'' in the
paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
considering the time taken to verify different candidate set instances.  The claim that
$T_{\mathit{iso\_test}}$ does not change much is not justified experimentally, and no consideration
is given as to whether this would hold true for other subgraph matching algorithms.

Moving forwards, the (simpler form of the) query response time equation is repeated by
\citet{DBLP:conf/vldb/ZhaoYY07}. Again, the work has a focus on reducing the candidate set size
through indexing. The authors appear to believe that the cost of the isomorphism test is not a major
factor in influencing the result, and focus on the remaining terms in the equation. They use the
``average cost'' of a subgraph isomorphism test as a constant, without considering that the average
cost could be influenced by the character of the candidate set.

The equation is also used by \citet{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
verification time dominates the Query Response Time [s]ince the computational complexity of
$T_{\mathit{iso\_test}}$ is NP-Complete''. They note that \begin{displayquote}``Approximately, the
    value of $T_{\mathit{iso\_test}}$ does not change too much with the difference of query. Thus,
the key to reducing query response time is to minimize the size of the candidate answer
set''.\end{displayquote} This claim becomes understandable when one examines the subgraph matching
algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
predates techniques like arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
pattern but only one in the target\footnote{Although interestingly, this algorithm effectively does
forward-checking and has a variable-ordering heuristic, before these concepts appeared in the
constraints literature.}.

Without using the equation, \citet{DBLP:conf/sigmod/ChengKNL07} state that since subgraph
isomorphism is NP-complete, processing by a sequential scan is infeasible. They introduce new
filtering techniques to try to avoid the subgraph isomorphism step. A similar claim is made by
\citet{DBLP:conf/icde/ZhangHY07} in an introduction of another indexing technique: ``obviously it is
inefficient to perform a sequential scan on every graph in the database, because the subgraph
isomorphism test is expensive''.

Muddying the waters slightly, a survey by \citet{DBLP:journals/datamine/HanCXY07} states that
``large volumes of data'' (not \NP-completeness) is the reason for using indexing in these systems.
However, in a description of a system tailored to biological networks,
\citet{DBLP:conf/edbt/ZhangLY09} state that \begin{displayquote}``Since the size of the raw database
    graph is small, it can be easily fit in the main memory.  However, the query (matching) time
will be very long due to the NP-hard complexity.''\end{displayquote} They suggest that indexing is
a way of avoiding this.

\citet{DBLP:conf/icdcs/CaoYWRL11} propose a privacy-preserving cloud graph database system using
filter~/ verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
infeasible to employ such a costly solution'' which simply checks every graph for a match.
\citet{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
subgraph isomorphism is an \NP-complete problem, a filter-and-verification method is usually
employed to speed up the search efficiency of graph similarity matching over a graph set''.
Similarly, after reviewing the literature, \citet{DBLP:journals/vldb/YuanM13} argue that subgraph
querying is costly because it is \NP-complete, and that indices can improve the performance of graph
database queries. Again, new indexing techniques are introduced. Subsequently, \citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
filtering techniques. They state that performing a query against each graph in the dataset
``obviously does not scale, as subgraph isomorphism is \NP-complete''. In describing a system which
returns a special subset of graphs which match a query, \citet{DBLP:journals/tkde/ZhengLZHZ16}
suggest that it is ``NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
``in order to improve the time efficiency'' they use an indexing system to ``avoid as many costly
subgraph isomorphism checkings as possible''. Their index takes tens of thousands of seconds to
build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
And \citet{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
subgraph isomorphism algorithm, such as Ullmann or VF2'', and argue that ``in order to speed up
query processing'', they need to create indices.

The filter~/ verify paradigm also influences other research. For example, it is used by
\citet{DBLP:conf/icde/TianP08} in an approximate subgraph searching system: they state that
Ullmann's algorithm ``is prohibitively expensive for querying against [a] database with a large
number of graphs'', and that indices are used ``to filter out graphs that do not match the query''.
More recently, \citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
using a filtering step to avoid subgraph isomorphism calls (in the opposite direction, so queries
are now target graphs). \citet{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
matching with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
are ``usually costly for large graphs'' because they ``do not utilize any index structure''. They
propose an indexing structure, which takes over 2,000 seconds to construct, and uses nearly 2GBytes
of space.  There is also research into maintaining indices when the set of target graph changes: for
example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices. And in
describing a system for reusing results of queries which are sub- or super-graphs of previous
queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
entails the \NP-complete problem of subgraph isomorphism'', and place ``an emphasis on the number of
unnecessary subgraph isomorphism tests''.

After some early ambiguity, then, it becomes clear that the intent behind filter~/ verify systems is
to reduce the number of subgraph isomorphism calls, and that the cost of loading graphs from disk is
not considered to be problematic. It is worth noting that the entire test datasets from most of
these papers will comfortably fit in RAM on a modern desktop machine, even when an adjacency matrix
representation is used.

Thus we can see that there are two critical beliefs underlying all of this work---firstly, that
subgraph isomorphism is necessarily hard because it is \NP-complete, and secondly, that there are
ways of identifying unsatisfiable instances using short proofs that a subgraph isomorphism algorithm
will not detect, but that an indexing system can. Throughout, the cost models used assume that the
time for a subgraph isomorphism query does not particularly depend upon the instance, and nowhere is
it considered that a good subgraph isomorphism algorithm should be able to eliminate
obviously-unsatisfiable instances with a similar time requirement to an indexing system.

These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
will immediately detect if a pattern contains two red vertices, whilst the target graph contains
only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
note, some variation on VF2 \citep{DBLP:journals/pami/CordellaFSV04} is the usual matching algorithm of choice for
graph database systems, although \citeauthor{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
chosen. Other approaches have been considered, albeit not with algorithms strong enough to
establish all-difference. For example, \citet{DBLP:journals/pvldb/ShangZLY08}
propose an algorithm which makes use of the frequency of various features to guide search;
\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
effective, even on families of graphs for which it was not designed.

We therefore believe it would be unlikely to cause too much astonishment if we suggested that a
better matching algorithm could be dropped in as a black box replacement in graph databases systems
to improve their performance.  This is not our claim. Instead, this paper shows that better
algorithms invalidate the flawed premise underlying the entire filter~/ verify approach. We will now
demonstrate that filter~/ verify is simply a poor workaround for the kinds of deficiency in VF2
demonstrated towards the end of the previous section.

\subsection{Is Filtering Necessary?}

To demonstrate that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
preprocessing, we look at four datasets commonly used in graph indexing evaluations
\citep{o:10.1371/journal.pone.0076911}. We do not claim that these are high-quality datasets or
that the associated queries are sensible, merely that they are widely used.

\begin{itemize}
    \item The AIDS dataset contains graphs representing 40,000 chemical molecules. These graphs are
        labelled, and are fairly small (mean 45 vertices) and sparse. Following
        \citeauthor{o:10.1371/journal.pone.0076911}, the queries are compounds with 8, 16 or 32
        edges.
    \item The PDBS dataset \citep{o:HeLCBBSKMR02} contains 30 labelled graphs representing DNA, RNA,
        and proteins, each duplicated twenty times. These
        can be relatively large, having up to tens of thousands of vertices, but are extremely
        sparse. The queries are randomly selected connected subgraphs and do not have a real-world
        meaning.
    \item The PCM dataset \citep{o:doi:10.1093/bioinformatics/btr163} contains 50 protein contact
        maps, each duplicated four times. These graphs have under a thousand vertices and below
        twenty thousand edges; they are slightly less sparse than the previous two datasets. As for
        PDBS, the queries are randomly generated and do not have a particular meaning.
    \item The PPI dataset contains 20 protein interaction networks, with up to ten thousand
        vertices. The queries have either four or eight vertices.
\end{itemize}

\begin{figure}[t]
    \centering
    \hspace*{1em}
    \includegraphics*{plots/biiiig-data-aids.pdf}
    \hfill
    \includegraphics*{plots/biiiig-data-pcms.pdf}
    \hspace*{1em}

    \vspace*{1em}

    \centering
    \hspace*{1em}
    \includegraphics*{plots/biiiig-data-pdbs.pdf}
    \hfill
    \includegraphics*{plots/biiiig-data-ppigo.pdf}
    \hspace*{1em}

    \caption{Cumulative number of instances solved as a function of search space size, using four
    graph database datasets and a simple domain-based subgraph isomorphism algorithm. Note that the
    $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
    \label{figure:biiiig-data}
\end{figure}

Rather than the modern techniques discussed used by Glasgow and LAD, we deliberately select a very
simple starting point: a forward-checking algorithm which uses the smallest domain first heuristic,
the simple counting all-different propagator, and degree and neighbourhood degree sequence filtering
at the top of search. Even this simple algorithm finds every instance nearly trivial, and no
exponential behaviour is observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest}
instance for the AIDS dataset requires 59 recursive calls, and 146,175 of the instances can be
solved without search (we measure recursive calls to demonstrate hardness, rather than algorithm
tuning and good constant factors; $10^4$ to $10^6$ recursive calls per second is feasible).  For
PDBS, the hardest problem requires 1,776 recursive calls (which occurs 20 times, due to duplicated
queries), and none of the rest require more than 182 calls. For PCM the hardest problem requires
3,261 calls, and for PPI, 26. In other words, none of these instances are in any way computationally
hard even for a simple domain-based algorithm, even before introducing stronger filtering,
inference, or heuristics. Furthermore it is certainly not the case that a sequential scan is
infeasible as so many filter~/ verify papers claim.  The exponential behaviour seen by others in
these instances is purely down to a bad choice of subgraph isomorphism algorithm, and these
instances are not inherently ``really hard''.

There are at least four benefits beyond simplicity towards abolishing filter~/ verify:

\begin{itemize}
    \item Domain filtering is useful on both satisfiable and unsatisfiable instances, whilst
        filtering can only eliminate trivially unsatisfiable instances, and is entirely wasted on
        satisfiable instances. Domain filtering is also useful even on relatively hard instances,
        since the information cuts down the search space rather than providing a simple ``yes'' or
        ``no''.

    \item Domain-based heuristics are much stronger than VF2's adjacency branching rules. Picking
        from small domains dynamically allows search to focus on the hardest part of the problem.

    \item This approach automatically combines features. For example, a pattern~/ target pair may
        have matching label features, and matching degree features, but if the only red pattern
        vertex has degree three whilst no red target vertex has degree more than two, domain
        filtering will detect this immediately. When combined with all-different propagation and
        maintained during search, this effect is even stronger.

    \item Finally, as indexing systems get more and more complex in an attempt to filter out a few more
        instances where VF2 performs poorly, the cost of index construction and maintenance is
        considerable.  Indices proposed in the literature often take many hours to build, and can
        consume much more space than the original graphs.
\end{itemize}

In other words, using domain-based algorithms would not simply be a viable alternative to filter~/
verify with VF2, but rather would be a much better solution.

\subsection{Rethinking Graph Matching for Database Systems}

The evidence so far clearly shows a need for graph database systems to be redesigned making use of
knowledge from constraint programming and artificial intelligence, and that this knowledge must inform
the systems as a whole, and not just parts of them.  Having demonstrated that filter~/ verify is
both theoretically and practically a flawed approach to designing graph databases based upon a
repeated misconception that \NP-completeness means long runtimes, we now briefly discuss how future
graph database systems should be designed, and how such systems can help support strong subgraph
isomorphism algorithms.

First, though, we briefly consider whether other lessons can be learned from constraint programming:
unfortunately indexing is not the only incorrect design choice being in these systems. For example,
recent work by \citet{DBLP:conf/edbt/KatsarouNT17} looks at what they call ``straggler'' queries,
which are the small subset of queries that they observe taking much longer than others to solve. The
authors present an approach to address this perceived problem that they describe as novel: ``instead
of trying to come up with new algorithms for sub-iso testing, we utilize isomorphic query rewritings
and existing alternative algorithms in parallel''. Their first claim is that permuting graphs before
running the subgraph isomorphism algorithm can lead to ``wildly different'' execution times. They
note that VF2 does not define ``a strict order in which the nodes of the query are matched'', and so
permuting the graphs (for example, by using degree or label frequency information) before search can
sometimes give orders of magnitude improvements. The connection to variable- and value-ordering
heuristics is not noted, and no consideration is given to even the simplest dynamic ordering
heuristics like smallest domain first \citep{DBLP:journals/ai/HaralickE80}. A little thought shows
that of the orderings proposed, those involving placing rare labels first are effectively poor
approximations to a static ``smallest domain at top of search first'' ordering, whilst the remainder
use degree as we did earlier in this paper. However, \citeauthor{DBLP:conf/edbt/KatsarouNT17} do
not investigate any algorithm which employs domains, let alone strong variable- or value-ordering
heuristics or all-different propagation, and do not consider that their apparent successes could be
due to VF2's weaknesses rather than an inherent property of \NP-completeness. We saw in the previous
section that although permuting input graphs could improve VF2's behaviour on some instances, doing
so does not make its performance come close to that of domain-based algorithms.
\citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up new research
directions, striving to find appropriate, per-query, isomorphic rewritings, in combination with
alternate per-query sub-iso algorithms that can yield large improvements''. We believe it is
important instead to emphasis previous research directions that have already solved most of this
problem, before any further research effort is wasted.

\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different algorithms find different
queries hard''. To address this, they run many subgraph isomorphism algorithms and input
permutations in parallel; the extensive literature on parallel portfolios in general
\citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for subgraph
isomorphism in particular, is not noted (and nor are portfolios mentioned when they say that ``using
machine learning models to predict which version \ldots to employ per query is of high interest'').
The evidence so far in this paper suggests that we should be wary. It is certainly true that there
are some instances that all algorithms find hard, and there are good theoretical reasons to believe
that these instances genuinely are really hard.  Furthermore, algorithm portfolios are also
well-known to be a successful technique, and are beneficial even with modern subgraph isomorphism
algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there
are many instances that VF2 finds hard that should not be hard, and that are not hard for other
algorithms.  If permuting graphs sometimes addresses the difficulty of some of these instances, then
it is likely that they are not genuinely hard instances at all, and perhaps it would be better
simply to start using an algorithm with domain tracking and domain-based ordering heuristics, rather
than assembling a portfolio of bad algorithms in the hopes that at least one of them will often get
lucky.

To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
this section. What if we had not used a smallest domain first heuristic, and instead used only
degree-based heuristics explained earlier in this paper? Although we still find all four of these
datasets easy, and the changes to the runtimes are hard to detect (more time is spent in
initialisation than in search), the worst case in the AIDS dataset now takes 4,051 recursive calls
as opposed to 59.  How about if, additionally, we do not detect domain wipeouts until attempting to
branch on an empty variable?  Suddenly three of the four datasets become extremely hard, with many
instances now timing out after making hundreds of millions of recursive calls; the PPI dataset now
requires up to 168,451 calls for some queries. Furthermore, slight changes to the input vertex
ordering can now make some of these ``hard'' instances easy again.

This backs up our suspicions: the apparent success of \citet{DBLP:conf/edbt/KatsarouNT17}'s
technique is due to the ease of sometimes getting better results out of a poor algorithm, rather
than being an inherent hardness property, and it would not help on really hard problem instances. We
point to \citet{DBLP:journals/heuristics/Gent98}'s admonition on how heuristics are benchmarked as
being similarly relevant here: \begin{displayquote}``Benchmark problems should be hard. I report on
    the solution of the five open benchmark problems introduced \ldots for testing bin packing
    problems. Since the solutions were found either by hand or by using very simple heuristic
    methods, these problems would appear to be easy. In four cases I give improved packings to
    refute conjectures that previously reported packings were optimal, and I give a proof that the
    fifth conjecture was correct. \ldots Future experimenters should be careful to perform tests on
problems that can reasonably be regarded as hard.''\end{displayquote} Claiming huge improvements
from an algorithm portfolio consisting only of variations of poor algorithms does not demonstrate a
genuine improvement over the state of the art.

We have seen, then, that domain tracking and domain-based ordering heuristics are critical, and we
have seen that these alone are enough to make indexing and other techniques irrelevant. There
is therefore no point in continued research such as that by
\citet{DBLP:conf/gbrpr/CarlettiFV15,o:Carletti16}
into tweaking VF2's connectivity-based heuristics.  However, we do not claim that the ultimate
``big data'' subgraph matching algorithm already exists: on the contrary, there is likely to be
plenty of room for future improvement now that we understand the importance of getting algorithm
design right.  For example, a major disadvantage of using domains is the relatively expensive
initialisation costs, which quickly add up when dealing with large numbers of trivial instances.
Employing a presolver is an obvious approach---and VF2 is actually good in this role
\citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities. For example,
minimal or lazy forward-checking
\citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
constructing every domain upfront, although adopting this require alternatives to smallest domain first
and to all-different. Such an approach may also be beneficial for huge target graphs, where having
domains range over the entire target is impractical.

There is also scope for precalculating supporting information about target graphs. For example,
neighbourhood degree sequences and supplemental graphs could both be pre-computed and stored. The aim
here is to reduce the initialisation costs of a good subgraph isomorphism algorithm, and not to
provide indexing (although additionally using this information as an index may not hurt, if
initialisation is still costly).

Our central message, though, is not just to say that graph database systems must use a better
algorithm.  Instead, this paper shows that such systems need to be designed with a better
understanding of the empirical hardness of \NP-complete problems, that subgraph isomorphism
algorithms must not be treated as a black box, and that lessons learned in constraint programming
and artificial intelligence must not go unheeded in other domains.

\section{Conclusion}

We have shown how to generate small but hard instances for the non-induced and induced subgraph
isomorphism problems, which will help offset the bias in existing datasets. For non-induced
isomorphisms, behaviour was as in many other hard problems, but for induced isomorphisms we
uncovered several interesting phenomena: there are hard instances far from a phase transition,
constrainedness predicts this, and existing general techniques for designing heuristics do not work
in certain portions of the parameter space.

When labels were introduced, we saw that VF2 finds some instances hard which other algorithms find
easy. We looked at this kind of instance in more detail, and argued that this was due to flaws in
VF2's design, rather than an interesting property of \NP-completeness. Although inevitably there
will be instances that every algorithm finds hard, we see no excuse for an algorithm to exhibit
exponential behaviour in the case when there are two red vertices in a pattern and only one in a
target.

Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, most
notably being the misdesign of graph database systems. By not trying even the simplest constraint
programming techniques from the literature, and ignoring all that is known about the empirical
hardness of \NP-complete problems, members of the graph databases community have misled themselves
into believing that extensive research into supporting techniques is important, whereas really all
they are doing is working around some but not all of the defects in their choices of subgraph
isomorphism algorithms. With this new understanding of what does and does not make subgraph
isomorphism hard, it is time for a radical rethink of how graph database systems work.

\bibliography{paper}
\bibliographystyle{theapa}

\clearpage
\appendix
\section{Reproducing These Experiments}

\end{document}